
```{r}
install.packages("rlang")
```

```{r}
# w  install.packages("tidyverse")
library(tidyverse)
library(dplyr) 
library(ggplot2)
data1 <- read.csv("C:/Users/Others/Desktop/updated data.csv") 
head(data1)
```


```{r}
summary(data1)
```
Data Preprocessing

```{r}
data2 <- data1 %>%
mutate(exang = if_else(exang == 1, "YES" ,"NO"), cp = if_else(cp == 1, "ATYPICAL ANGINA",
if_else(cp == 2, "NON-ANGINAL PAIN", "ASYMPTOMATIC")),
restecg = if_else(restecg == 0, "NORMAL",
if_else(restecg == 1, "ABNORMALITY", "PROBABLE OR
DEFINITE")),
slope = as.factor(slope), ca = as.factor(ca),
thal = as.factor(thal),
target = if_else(target == 0, "NO",
                if_else(target == 1, "YES","YES")),
) %>%
mutate_if(is.character, as.factor) %>%
dplyr::select(target, sex, trestbps, exang, cp, restecg, slope, ca, thal,everything())

head(data2)

```
Data Exploration
```{r}
ggplot(data2, aes(x=data2$target, fill=data2$target))+ geom_bar()+
xlab("Heart Disease")+ ylab("count")+
ggtitle("Presence & Absence of Heart Disease")+
scale_fill_discrete(name= 'Heart Disease', labels =c("Absence", "Presence"
))

```

Frequency of values of age
```{r}
data2 %>%
group_by(age) %>% count() %>% filter(n > 10) %>% ggplot() +
geom_col(aes(age, n), fill = 'orange') + ggtitle("Age Analysis")+
xlab("Age")+ ylab("Age Count")

```
By observing the graph we can say that the dataset contains more people who are in the range of age 55 to 60.

Compare blood pressure across the chest pain

```{r}
data2 %>%
ggplot(aes(x=sex, y=trestbps)) + geom_boxplot(fill = 'green') + xlab("Sex") +
ylab("Blood Pressure") + facet_grid(~cp)

```
From above graph we can say that Females are experiencing highest blood pressure when they are having Asympomatic chest pain.
```{r}
data2 %>%
ggplot(aes(x=sex, y=chol)) + geom_boxplot(fill = 'orange') + xlab("Gender") + ylab("Cholesterol") + facet_grid(~cp)
```
From above graph we can say that Females are experiencing high Cholestrol levels when they are having Asympomatic chest pain.

```{r}
data2 %>%
ggplot(aes(x=sex, y=trestbps))+ geom_boxplot(fill ='purple')+ xlab('sex')+
ylab('BP')+ facet_grid(~cp)

```

```{r}
data2 %>%
ggplot(aes(x=sex, y=chol))+ geom_boxplot(fill ='orange')+ xlab('sex')+
ylab('Chol')+ facet_grid(~cp)

```
## Correlation

```{r}
library(corrplot)
```
```{r}
cor_heart <- cor(data2[, 10:14]) 
cor_heart
```
```{r}
corrplot(cor_heart, method ='square', type='upper')
```
Checking the null values in the dataset.

```{r}
is.null(data2)
```
```{r}
library(lattice) 
library(caret)
```

Now let us divide the data in the test and train set.
In this project, I have divided the data into an 70: 30 ratio. That is, the training size is 70% and testing size is 30% of the whole data.


```{r}
set.seed(10)
inTrainRows <- createDataPartition(data2$target,p=0.7,list=FALSE)
trainData <- data2[inTrainRows,] 
testData <- data2[-inTrainRows,]
nrow(trainData)/(nrow(testData)+nrow(trainData))

```
Classifier

a) SVM model
Using the Cross Validation method for Classification.

```{r}
install.packages("kernlab")
```

```{r}
grid <- expand.grid(C = 10^seq(-5,2,0.5))
train_control = trainControl(method = "cv", number = 10)
svm_grid <- train(target ~., data = trainData, method = "svmLinear", trControl = train_control, tuneGrid = grid)
svm_grid

```
Confusion Matrix to predict the accuracy of SVM

```{r}
pred_tree1 <- predict(svm_grid, testData) 
confusionMatrix(testData$target, pred_tree1)
```

Using the classifier of Decision tree to classify the data.
```{r}
library(bitops)
library(rattle)
library(rpart)
```

   

```{r}
train_control = trainControl(method = "cv", number = 10) 
hypers = rpart.control(minsplit = 12, maxdepth = 7,minbucket = 10)

tree1 <- train(target~ .,control = hypers, data = trainData,method = "rpart1SE",trControl = train_control)
tree1

```

Getting errors in decison tree. I have to check again

Clustering

Creating dummy variables

```{r}
heart_norm<- data2 %>% select(-c(target))
preprocess <- preProcess(heart_norm, method=c("center","scale")) 
norm <- predict(preprocess, heart_norm)
summary(norm)

```
```{r}
heart_norm$class<-data2$class
```

```{r}
norm_dummy <- dummyVars(~ ., data = norm)
norm_dummies <- as.data.frame(predict(norm_dummy, newdata = norm)) 
head(norm_dummies)

```
```{r}
heart_dummies <- as.data.frame(norm_dummies) 
heart_dummies <- na.omit(heart_dummies) 
head(heart_dummies)
```
```{r}
library(caret)
library(ggplot2)
```





```{r}
#install.packages("factoextra")
#install.packages("caret")
#install.packages("ggplot2")
#library(caret)
update.packages("ggplot2")
update.packages("caret")
library(ggplot2)
library(caret)
library(factoextra)
library(cluster)
library(lattice)
set.seed(13)
heart2<-heart_dummies
preprocess<-preProcess(heart2,method=c("center","scale")) 
pred<-predict(preprocess,heart2)

```

finding optimal no.of clusters
```{r}
fviz_nbclust(heart2, kmeans, method = "wss")
```

```{r}
fviz_nbclust(heart2, kmeans, method = "silhouette")
```
```{r}
gap_stat <- clusGap(heart2,FUN = kmeans, nstart = 25,K.max = 10,B = 50)

```
```{r}
fviz_gap_stat(gap_stat)
```
Choosing K=3

```{r}
km <- kmeans(heart2, centers = 3, nstart = 25) 
km
```

Visualizing the cluster

```{r}
fviz_cluster(km, data = heart2)
```
```{r}
# Calculate PCA 
pca = prcomp(pred) 
# Save as dataframe
rotated_data = as.data.frame(pca$x)
# Add original labels as a reference 
rotated_data$Clusters = as.factor(km$cluster) # Plot and color by labels
ggplot(data = rotated_data, aes(x = PC2, y = PC1, col = Clusters)) + geom_point()


```

